Ecosystem size and complexity dictate riverine biodiversity

### 2021-01-08 manuscript.Rmd
Update manuscript.Rmd

---
title: "Ecosystem size and complexity dictate riverine biodiversity"
author:
- Akira Terui^[Department of Biology, University of North Carolina at Greensboro]
- Seoghyn Kim\footnotemark[1]
- Christine L. Dolph^[Department of Ecology, Evolution and Behavior]
output:
  html_notebook: default
  word_document: default
abstract: |
  |This is the abstract.
bibliography: reference.bib
csl: nature.csl
---

@@

## read gis files
filename <- list.files(path = normalizePath('../empirical/data_gis'), full.names = T)
wsd_subset <- lapply(filename[str_detect(filename, "wsd_subset")], st_read)
point_subset <- lapply(filename[str_detect(filename, "point_subset")], st_read)
shape <- lapply(filename[str_detect(filename, "shape")], st_read)
shape[[1]] <- st_set_crs(shape[[1]], st_crs(wsd_subset[[1]]))

## read theory file
dat_sim <- read_csv(normalizePath('../theory/result/result_sim2020-11-10.csv'))

@@

source(normalizePath('../empirical/analysis_lm_bf.R'))

@@

Ecologists have long sought to understand the general drivers of biodiversity. One of the most robust empirical generalizations in ecology is the positive relationship between species richness and area, i.e., the species-area relationship (the SAR)[@lomolinoEcologyMostGeneral2000a]. In 1921, Arrhenius[@arrheniusSpeciesArea1921] formulated the SAR as a power-law $S = cA^z$ ($S$ is the number of species observed in a given geographic area $A$), an equation currently known as the Arrhenius species-area relationship. Since then, the spatial scaling of species richness has been observed in nearly all taxonomic groups[@triantisIslandSpeciesArea2012], although the underlying mechanisms are still hotly debated[@ben-hurDisentanglingMechanismsUnderlying2020]. The SAR is ubiquitous because multiple mechanisms produce an apparently similar pattern. Larger ecosystems typically support more diverse metacommunities due to increased habitat diversity[@kallimanisHowDoesHabitat2008a], larger metacommunity size[@ben-hurDisentanglingMechanismsUnderlying2020;@hubbellUnifiedNeutralTheory2001], and/or enhanced colonization dynamics[@lososTheoryIslandBiogeography2009]. Importantly, the SAR provides the foundation for global conservation efforts. For example, conservation ecologists have used the SAR estimates to design marine and terrestrial protected areas, which currently encompass more than 30 million km² globally.

Many ecosystems, however, possess complex spatial structure that cannot be represented by area - a dimension referred to as the scale-invariant complexity[@rinaldoEvolutionSelectionRiver2014; @rodriguez-iturbeFractalRiverBasins2001]. Such complexity is evident in branching ecosystems, including rivers, trees, and mountain ranges, to name just a few[@rodriguez-iturbeFractalRiverBasins2001]. Geomorphic or biological processes generate a pronounced self-similarity in complex branching patterns such that the part and the whole look alike[@rodriguez-iturbeFractalRiverBasins2001]. Even though the branching structure is independent of spatial scale, it forms a physical template that dictates habitat diversity and dispersal corridors for living organisms[@grantLivingBranchesPopulation2007; @rinaldoRiverNetworksEcological2018;@tonkinRoleDispersalRiver2018]. Limited, but accumulating evidence suggests that classical concepts cannot predict ecological dynamics driven by branching structure[@faganConnectivityFragmentationExtinction2002; @grantUseMultipleDispersal2010; @teruiMetapopulationStabilityBranching2018], and this recognition has led to recent developments of spatial theories devoted to complex branching ecosystems[@altermattDiversityRiverineMetacommunities2013; @carraraComplexInteractionDendritic2014; @carraraDendriticConnectivityControls2012; @teruiMetapopulationStabilityBranching2018]. However, most research either has explored the consequences of branching complexity for local variation in species richness or relied on theoretical arguments with limited replicates of artificial branching networks. At present, we lack a comprehensive evaluation of how branching complexity, in combination with ecosystem size, scales biodiversity at the metacommunity level. Filling this knowledge gap may provide common ground for achieving successful conservation in spatially complex ecosystems, where the accelerated species loss threatens the delivery of ecosystem services.

Here, we hypothesize that ecosystem size and complexity dictate biodiversity patterns in rivers, a prime example of complex branching ecosystems.
The recurrent merging of diverse streams and rivers produces larger, ecologically distinct channels, ultimately forming a fluvial network with fractal branching patterns[@rodriguez-iturbeFractalRiverBasins2001]. As such, the complexity of branching structure, which we define here as the probability of branching per unit river distance [@teruiMetapopulationStabilityBranching2018; @yeakelSynchronisationStabilityRiver2014a], should control the habitat heterogeneity of the ecosystem (habitat diversity per unit area). Meanwhile, ecosystem size (watershed area) should determine the metacommunity size and total habitat diversity (area x heterogeneity). We predict that ecosystem size and branching complexity increase $\gamma$ diversity by enhancing either $\alpha$ and/or $\beta$ diversity under different ecological scenarios. The present study combines theory and statistical analysis of extensive community data to provide crucial insights into how ecological communities are structured in complex branching networks.

First, we theoretically analyzed the influences of ecosystem size and branching complexity on $\gamma$ diversity. We constructed a general simulation model of metacommunity dynamics [@thompsonProcessbasedMetacommunityFramework2020], in which 50 competing species disperse and spatially interact in a random branching network. Simulated species with different niche in abiotic conditions (optimum and width) compete for resources with varied strengths. For simplicity, other ecological traits, such as dispersal probability, are assumed to be identical among species. Here, we depicted branching ecosystems as a spatial network of connected habitat patches (or a discretized river section; see **Figure 1**) where mean environmental conditions at each habitat patch are determined by two factors: (1) the headwater environments (the most upstream habitat patch) and (2) local environmental noise. Environmental values at the headwaters are drawn randomly from a normal distribution and propagates downstream with local environmental noise (i.e., the spatial autoregressive process with white noise). These environmental values recurrently 'mix' at confluences considering the relative size of joining tributaries (see Methods). Therefore, our simulation resembles natural processes of how branching river networks create diverse habitats in a metacommunity. We used `r n_scenario` sets of parameter combinations with varied ecological (dispersal and competition strength) and landscape scenarios (the environmental variation at headwaters and the degree of local environmental noise). Under each scenario, we simulated metacommunity dynamics in `r max(dat_sim$n_rep)` branching networks that have different ecosystem size (the number of habitat patches: `r min(dat_sim$n_patch)` to `r max(dat_sim$n_patch)`) and complexity (branching probability: `r round(min(dat_sim$p_branch), 2)` to `r round(max(dat_sim$p_branch), 2)`) to examine their effects on species diversity.

Our theoretical analysis yielded results consistent with our prediction. Ecosystem size and complexity both increased $\gamma$ diversity under a realistic landscape scenario (**Figures 2 and 3**), where the environmental variation at headwaters ($\sigma_h$) is greater than the degree of local environmental noise ($\sigma_l$) [@petersonModellingDendriticEcological2013; @teruiMetapopulationStabilityBranching2018]. The relationships had a characteristic of power-law (i.e., linear in a log-log scale) and were consistent under various ecological scenarios. The strength of competition and dispersal processes (dispersal distance and probability) did not change the form of the relationships between $\gamma$ diversity and ecosystem properties. Hence, the ecosystem size and complexity are both expected to increase $\gamma$ diversity regardless of ecological scenarios.

Dispersal processes, however, affected underlying mechanisms that produce the positive effects of ecosystem size and complexity on $\gamma$ diversity. We observed a greater contribution of $\beta$ diversity (defined as $\frac{\gamma}{\alpha}$) to increased $\gamma$ diversity when dispersal limitation exists (i.e., species travel short distances). This pattern reflects significant spatial turnover of species composition over the branching network. In contrast, once the dispersal limitation was relaxed (species travel long distances), a clear increase in $\alpha$ diversity underpinned the positive relationships between $\gamma$ diversity and ecosystem properties. The results agree with previous predictions that increased dispersal homogenizes community composition while enhancing local diversity through increased immigrants from suitable habitat patches (i.e., mass effects)[@thompsonProcessbasedMetacommunityFramework2020]. These patterns were consistent across different levels of dispersal probabilities (**cite SI figure**). The strength of competition decreased maximum levels of $\alpha$ diversity but did not change the functional forms of the scaling relationships with ecosystem properties (**Figures 2 and 3**). In summary, our theory highlights how the apparently similar patterns in $\gamma$ diversity emerges through different ecological pathways.

Influences of ecosystem size and complexity differed significantly in their dependence on landscape scenarios. Ecosystem size had positive effects on $\gamma$ diversity regardless of landscape scenarios, although the slopes were steeper with greater environmental variation (higher $\sigma_h$ and/or $\sigma_l$). This result is attributable to the fact that larger ecosystems can hold more individuals in a metacommunity[@hubbellUnifiedNeutralTheory2001]. In contrast, we observed limited or no influences of branching complexity when local environmental noise was equal to or exceeded environmental variation at headwaters ($\sigma_l \ge \sigma_h$). Under this scenario, having more branches may not contribute to the habitat diversity in the ecosystem because local environmental noise masks environmental differences between branches. Therefore, this landscape scenario decouples the intimate relationship between branching structure and habitat diversity, thereby eliminating the positive effect of branching complexity on $\gamma$ diversity. This theoretical prediction may not apply to pristine to semi-natural river networks where individual streams with diverse geological and/or climatic backgrounds show distinct environmental conditions, including water temperature, water chemistry, substrate, and flow/sediment regimes. Instead, it may be more relevant to severely altered landscapes where human disturbance compromises the environmental distinctiveness of each branch through, for example, flow regulations by dams. Hence, our theory has important implications for riverine biodiversity conservation by pointing the crucial role of habitat diversity produced by branching structure[@holtRoleBranchingMaintenance2018].

@@

Similarly, we found weak to moderate supports for the global models of $\alpha$ and $\beta$ diversity (Bayes factor: `r round(fit$alpha$bayes_factor, 1)` and `r round(fit$beta$bayes_factor, 1)`, respectively). In both regions, $\beta$ diversity responded significantly to ecosystem size and complexity. In our simulations, this pattern has emerged under the scenarios with dispersal limitation, which elegantly match the previous observations of stream fish movement. Direct (mark-recapture) and indirect observations (e.g., genotyping) recurrently revealed restricted movement of stream fish, typically limited to several tens to hundreds of meters in distance[@comteFishDispersalFlowing2018; @radingerPatternsPredictorsFish2014; @rodriguezRestrictedMovementStream2002]. The reciprocal agreement of theoretical and empirical patterns provides indirect but convincing evidence that dispersal limitation plays a key role in driving the associations between $\gamma$ diversity and ecosystem properties in rivers.

@@

The emerging complexity-diversity relationship points to several important avenues for riverine biodiversity conservation. First and foremost, there is now a clear need to explicitly consider the dimension of ecosystem complexity to achieve successful conservation. Human alterations, including flow regulation[@poffHomogenizationRegionalRiver2007], habitat fragmentation, and stream burial, may compromise or restrict access to the diverse habitats that complex branching networks may provide. Hence, it is imperative to recognize the role of branching complexity and minimize the homogenizing effects of human alterations. Second, the complexity perspective may provide viable options to reconcile conflicts between biodiversity conservation and societal needs. Although protected areas spearhead global conservation efforts, habitat preservation and restoration over large areas are often impractical because of conflicts with stakeholders. This is particularly true for riverine ecosystems, where numerous stakeholders are involved within a single watershed and enormous efforts are required to reach consensus[@palmerEcologicalRestorationStreams2014]. In such a scenario, spatial planning of small-scale conservation, such as restoration of connectivity to small tributaties, may be critical to enhancing the 'realized' network complexity that organisms may experience. Our theoretical framework may be used to design spatial arrangement of small-scale conservation that improves ecological outcomes at the metacommunity level.

While the prevailing evidence supports the importance of ecosystem size in scaling species diversity, ecosystem complexity - especially branching structure - has not received the attention it deserves. Ecosystems are inherently complex in its spatial structure, providing a physical template that creates a wide spectrum of niche opportunities for living organisms. Hence, our findings should be broadly applicable to many taxa and ecosystems. Generalization of our findings to other spatially complex ecosystems represents a frontier for future research.

@@

We estimated watershed characteristics (watershed area and branching probability), climatic variables (annual mean air temperature and cumulative precipitation), and land use (% forest). We delineated watershed polygons and riverlines using Hydrologically Adjusted Elevations of MERIT Hydro[@yamazakiMERITHydroHighResolution2019] with ArcMap 10.7 and SAGA 6.3.0 combined with an R package 'RSAGA'[@brenningRSAGASAGAGeoprocessing2018]. Watershed area was estimated based on the delineated watershed polygons using functions in an R package 'sf'[@pebesmaSimpleFeaturesStandardized2018]. Branching probability was estimated by fitting exponential distributions to histograms of branch length (unit: kilometer, river segment between successive confluences or a confluence and the outlet/upstream terminal) as $branch~length \sim Exp(\lambda)$. A derived statistical quantity $1 − e^{−\lambda}$ corresponds to the theoretical branching probability $P_b$. For each watershed, we estimated the spatial averages of air temperature and cumulative precipitation at a 1-km resolution, which were extracted from WorldClim 1.4[@hijmansVeryHighResolution2005] via R package 'raster'. We calculated % forest for each watershed using land use data in 2015 from Copernicus Global Land Service (100-m resolution)[@marcelbuchhornCopernicusGlobalLand2020].

@@

>>>>

---
title: "Ecosystem size and complexity dictate riverine biodiversity"
author:
- Akira Terui^[Department of Biology, University of North Carolina at Greensboro]
- Seoghyn Kim\footnotemark[1]
- Christine L. Dolph^[Department of Ecology, Evolution and Behavior, University of Minnesota]
- Taku Kadoya^[National Institute for Environmental Studies]
- Yusuke Miyazaki^[Shiraume Gakuen University]
- Jacques C. Finlay\footnotemark[2]
output:
  html_notebook: default
  word_document: default
abstract: |
  |This is the abstract.
bibliography: reference.bib
csl: nature.csl
---

@@

## read gis files
filename <- list.files(path = here::here('/empirical/data_gis'), full.names = T)
wsd_subset <- lapply(filename[str_detect(filename, "wsd_subset")], st_read)
point_subset <- lapply(filename[str_detect(filename, "point_subset")], st_read)
shape <- lapply(filename[str_detect(filename, "shape")], st_read)
shape[[1]] <- st_set_crs(shape[[1]], st_crs(wsd_subset[[1]]))

## read theory file
dat_sim <- read_csv(here::here('/theory/result/result_sim2020-11-10.csv'))

@@

source(here::here('/empirical/analysis_lm_bf.R'))

@@

Ecologists have long sought to understand the general drivers of biodiversity. One of the most robust empirical generalizations in ecology is the positive relationship between species richness and area, i.e., the species-area relationship (the SAR)[@lomolinoEcologyMostGeneral2000a]. In 1921, Arrhenius[@arrheniusSpeciesArea1921] formulated the SAR as a power-law $S = cA^z$ ($S$ is the number of species observed in a given geographic area $A$), an equation currently known as the Arrhenius species-area relationship. Since then, the spatial scaling of species richness has been observed in many taxonomic groups[@triantisIslandSpeciesArea2012], although the underlying mechanisms are still hotly debated[@ben-hurDisentanglingMechanismsUnderlying2020]. The SAR is ubiquitous because multiple mechanisms produce an apparently similar pattern. Larger ecosystems typically support more diverse metacommunities due to increased habitat diversity[@kallimanisHowDoesHabitat2008a], larger metacommunity size[@ben-hurDisentanglingMechanismsUnderlying2020;@hubbellUnifiedNeutralTheory2001], and/or enhanced colonization dynamics[@lososTheoryIslandBiogeography2009]. Importantly, the SAR provides the foundation for global conservation efforts. For example, conservation ecologists have used the SAR estimates to design marine and terrestrial protected areas, which currently encompass more than 30 million km² globally.

Many ecosystems, however, possess complex spatial structure that cannot be represented by area - a dimension referred to as the scale-invariant complexity[@rinaldoEvolutionSelectionRiver2014; @rodriguez-iturbeFractalRiverBasins2001]. Such complexity is evident in branching ecosystems, including rivers, trees, and mountain ranges, to name just a few[@rodriguez-iturbeFractalRiverBasins2001]. Geomorphic or biological processes generate a pronounced self-similarity in complex branching patterns such that the part and the whole look alike[@rodriguez-iturbeFractalRiverBasins2001]. Even though the branching structure is independent of spatial scale, it forms a physical template that dictates habitat diversity and dispersal corridors for living organisms[@grantLivingBranchesPopulation2007; @rinaldoRiverNetworksEcological2018;@tonkinRoleDispersalRiver2018]. Limited, but accumulating evidence suggests that classical concepts cannot predict ecological dynamics driven by branching structure[@faganConnectivityFragmentationExtinction2002; @grantUseMultipleDispersal2010; @teruiMetapopulationStabilityBranching2018], and this recognition has led to recent developments of spatial theories devoted to complex branching ecosystems[@altermattDiversityRiverineMetacommunities2013; @carraraComplexInteractionDendritic2014; @carraraDendriticConnectivityControls2012; @teruiMetapopulationStabilityBranching2018]. However, most research either has explored the consequences of branching complexity for local community structure[@altermattRiverNetworkProperties2013; @brownDendriticNetworkStructure2010] or relied solely on theoretical arguments[@holtRoleBranchingMaintenance2018]. At present, we lack a comprehensive evaluation of how branching complexity, in combination with ecosystem size, scales biodiversity at the metacommunity level. Filling this knowledge gap may provide common ground for achieving successful conservation in spatially complex ecosystems, where the accelerated species loss threatens the delivery of ecosystem services[@dudgeonFreshwaterBiodiversityImportance2006].

Here, we hypothesize that ecosystem size and complexity dictate biodiversity patterns in rivers, a prime example of complex branching ecosystems.
The recurrent merging of diverse streams and rivers produces larger, ecologically distinct channels, ultimately forming a fluvial network with fractal branching patterns[@rodriguez-iturbeFractalRiverBasins2001]. As such, the complexity of branching structure, which we define here as the probability of branching per unit river distance [@teruiMetapopulationStabilityBranching2018; @yeakelSynchronisationStabilityRiver2014a], should control the habitat heterogeneity of the ecosystem (habitat diversity per unit area). Meanwhile, ecosystem size (watershed area) should determine the metacommunity size and total habitat diversity (area x heterogeneity). We predict that ecosystem size and branching complexity increase $\gamma$ diversity by enhancing either $\alpha$ and/or $\beta$ diversity under different ecological scenarios. The present study combines theory and empirical analysis of extensive community data to provide crucial insights into how ecological communities are structured in complex branching networks.

First, we theoretically analyzed the influences of ecosystem size and branching complexity on $\gamma$ diversity. We constructed a general simulation model of metacommunity dynamics [@thompsonProcessbasedMetacommunityFramework2020], in which 50 competing species disperse and spatially interact in a random branching network. Simulated species with different niche in abiotic conditions (optimum and width) compete for resources with varied strengths. Here, we depicted branching ecosystems as a spatial network of connected habitat patches (**Figure 1**) where mean environmental conditions at each habitat patch are determined by two factors: (1) headwater environments (the most upstream habitat patch) and (2) local environmental noise. Environmental values at headwaters are drawn randomly from a normal distribution and propagate downstream with local environmental noise (i.e., a spatial autoregressive process with white noise). These environmental values recurrently 'mix' at confluences considering the relative size of joining tributaries (see Methods). Therefore, our simulation resembles natural processes of how branching river networks create diverse habitats in a metacommunity. We used `r n_scenario` sets of parameter combinations with varied ecological (dispersal and competition strength) and landscape scenarios (the environmental variation at headwaters and the degree of local environmental noise). Under each scenario, we simulated metacommunity dynamics in `r max(dat_sim$n_rep)` branching networks that have different ecosystem size (the number of habitat patches: `r min(dat_sim$n_patch)` to `r max(dat_sim$n_patch)`) and complexity (branching probability: `r round(min(dat_sim$p_branch), 2)` to `r round(max(dat_sim$p_branch), 2)`) to examine their effects on species diversity.

Our theoretical analysis yielded results consistent with our prediction. Ecosystem size and complexity both increased $\gamma$ diversity under a realistic landscape scenario (**Figures 2 and 3**), where the environmental variation at headwaters ($\sigma_h$) is greater than the degree of local environmental noise ($\sigma_l$) [@petersonModellingDendriticEcological2013; @teruiMetapopulationStabilityBranching2018]. The relationships had a characteristic of power-law (i.e., linear in a log-log scale) and were consistent under various ecological scenarios. The strength of competition and dispersal processes (dispersal distance and probability) did not change the form of the relationships between $\gamma$ diversity and ecosystem properties (compare panels in **Figures 2 and 3**). Hence, the ecosystem size and complexity are both expected to increase $\gamma$ diversity regardless of ecological scenarios.

Dispersal processes, however, affected mechanisms that underlie the positive effects of ecosystem size and complexity on $\gamma$ diversity (compare left and right columns in **Figures 2 and 3**). We observed a greater contribution of $\beta$ diversity (defined as $\frac{\gamma}{\alpha}$) to increased $\gamma$ diversity when dispersal limitation exists (i.e., species travel short distances). This pattern reflects significant spatial turnover of species composition over the branching network. In contrast, when the dispersal limitation was relaxed (species travel long distances), a clear increase in $\alpha$ diversity underpinned the positive relationships between $\gamma$ diversity and ecosystem properties. The results agree with previous predictions that increased dispersal homogenizes community composition while enhancing local diversity through increased immigrants from suitable habitat patches (i.e., mass effects)[@thompsonProcessbasedMetacommunityFramework2020]. These patterns were consistent across different levels of dispersal probabilities (**cite SI figure**). The strength of competition decreased maximum levels of $\alpha$ diversity but did not change the functional forms of the scaling relationships with ecosystem properties (**Figures 2 and 3**). In summary, our theory highlights how the apparently similar patterns in $\gamma$ diversity emerges through different ecological pathways.

Influences of ecosystem size and complexity differed significantly in their dependence on landscape scenarios. Ecosystem size had positive effects on $\gamma$ diversity regardless of landscape scenarios, although the slopes were steeper with greater environmental variation (higher $\sigma_h$ and/or $\sigma_l$) (). This result is attributable to the fact that larger ecosystems can hold more individuals in a metacommunity[@hubbellUnifiedNeutralTheory2001]. In contrast, we observed limited or no influences of branching complexity when local environmental noise was equal to or exceeded environmental variation at headwaters ($\sigma_l \ge \sigma_h$). Under this scenario, having more branches may not contribute to the overall habitat diversity in the ecosystem because local environmental noise masks environmental differences between branches. Therefore, this landscape scenario decouples the intimate relationship between branching structure and habitat diversity, thereby eliminating the positive effect of branching complexity on $\gamma$ diversity. This theoretical prediction may not apply to pristine or semi-natural river networks where individual streams with diverse geological and/or climatic backgrounds show distinct environmental conditions, including water temperature, water chemistry, substrate, and flow/sediment regimes[@altermattDiversityRiverineMetacommunities2013; @bendaNetworkDynamicsHypothesis2004; @mooreEmergentStabilityLarge2015; @nakamuraDisturbanceRegimesStream2000; @petersonModellingDendriticEcological2013; @teruiMetapopulationStabilityBranching2018]. Instead, it may be more relevant to severely altered landscapes where human disturbance compromises the environmental distinctiveness of branches through, for example, flow regulations by dams[@poffHomogenizationRegionalRiver2007]. Hence, our theory has important implications for riverine biodiversity conservation by pointing the crucial role of habitat diversity produced by branching structure.

@@

Similarly, we found weak to moderate supports for the global models of $\alpha$ and $\beta$ diversity (Bayes factor: `r round(fit$alpha$bayes_factor, 1)` and `r round(fit$beta$bayes_factor, 1)`, respectively). In both regions, $\beta$ diversity responded significantly to ecosystem size and complexity while $\alpha$ diversity showing a vague response to these ecosystem properties. In our simulations, this pattern has emerged under the scenarios with dispersal limitation, which elegantly match the previous observations of stream fish movement. Direct (mark-recapture) and indirect observations (e.g., genotyping) recurrently revealed restricted movement of stream fish, typically limited to several tens to hundreds of meters in distance[@comteFishDispersalFlowing2018; @radingerPatternsPredictorsFish2014; @rodriguezRestrictedMovementStream2002]. The reciprocal agreement of theoretical and empirical patterns provides indirect but convincing evidence that dispersal limitation plays a key role in driving the associations between $\gamma$ diversity and ecosystem properties in rivers.

@@

The emerging complexity-diversity relationship points to several important avenues for riverine biodiversity conservation. First and foremost, there is now a clear need to explicitly consider the dimension of ecosystem complexity to achieve successful conservation. Human-induced habitat alterations, including flow regulation[@poffHomogenizationRegionalRiver2007], habitat fragmentation[@fukushimaModellingEffectsDams2007; @moritaEffectsHabitatFragmentation2002], and stream burial, may compromise or restrict access to the diverse habitats that complex branching networks may provide. Hence, it is imperative to recognize the role of branching complexity and minimize the homogenizing effects of human activities. Second, the complexity perspective may provide insights into the spatial arrangement of conservation efforts. A growing body of evidence suggests the effectiveness of riverine reserves and local restoration, and how to arrange a spatial network of conservation sites in rivers is an area of active research[@koningNetworkGrassrootsReserves2020; @stollSmallImpoverishedRegional2013; @sundermannRiverRestorationSuccess2011]. The results of this study clearly indicate that riverine conservation networks should be designed to involve diverse tributaries, as inferred from a successful reserve network of the Salween basin in Thailand[@koningNetworkGrassrootsReserves2020].

While the prevailing evidence supports the importance of ecosystem size in scaling species diversity, ecosystem complexity - especially branching structure - has not received the attention it deserves. Ecosystems are inherently complex in its spatial structure, providing a physical template that creates a wide spectrum of niche opportunities for living organisms. Hence, our findings should be broadly applicable to many taxa and ecosystems. Generalization of our findings in spatially complex ecosystems represents a frontier for future research.

@@

We collected data for watershed characteristics (watershed area and branching probability), climatic variables (annual mean air temperature and cumulative precipitation), and land use (% forest). We delineated watershed polygons and riverlines using Hydrologically Adjusted Elevations of MERIT Hydro[@yamazakiMERITHydroHighResolution2019] with ArcMap 10.7 and SAGA 6.3.0 combined with an R package 'RSAGA'[@brenningRSAGASAGAGeoprocessing2018]. Watershed area was estimated based on the delineated watershed polygons using functions in an R package 'sf'[@pebesmaSimpleFeaturesStandardized2018]. Branching probability was estimated by fitting exponential distributions to histograms of branch length (unit: kilometer, river segment between successive confluences or a confluence and the outlet/upstream terminal) as $branch~length \sim Exp(\lambda)$. A derived statistical quantity $1 − e^{−\lambda}$ corresponds to the theoretical branching probability $P_b$. For each watershed, we estimated the spatial averages of air temperature and cumulative precipitation at a 1-km resolution, which were extracted from WorldClim 1.4[@hijmansVeryHighResolution2005] via R package 'raster'. We calculated % forest for each watershed using land use data in 2015 from Copernicus Global Land Service (100-m resolution)[@marcelbuchhornCopernicusGlobalLand2020].

@@

We developed linear models to examine the effects of macro-scale drivers on diversity metrics. Log-transformed diversity metrics ($\alpha$, $\beta$ and $\gamma$ diversity) at watershed $i$, $log_{10}~y_i$, are assumed to follow a normal distribution as $log_{10}~y_i \sim Normal(\mu_i, \sigma^2)$. The parameter $\mu_i$ was related to linear predictors, including log-transformed watershed area (ecosystem size; $log_{10}~WA$), log-branching probability (ecosystem complexity; $log_{10}~P_b$), region (coded as Hokkaido = 0, Midwest = 1; $region$), air temperature, precipitation, and proportion of forest land use (logit-transformed). Although our primary focus of this analysis is to assess the effects of watershed area and branching probability, we included climate and land use variables to statistically control these effects.

Climate and land use differed clearly between the regions: therefore, we used residuals of the simple linear regressions between either the climate or land use variable (response variable) and region (explanatory variable) to avoid correlation between explanatory variables. The residuals represent deviations from the regional averages; positive values indicate values greater than the regional average while negative values indicating values smaller than the regional average.

We constructed two models for each of diversity metrics: global and region-specific models. The global model assumes that the effects of watershed area and branching probability are consistent across biogeographic regions (i.e., no interactive effects of region with other linear predictors).

$$
\mu_i = \xi_0 + \xi_1~log_{10}~WA_i + \xi_2~log_{10}~P_{b,i} + \xi_3~region_i + \sum_k \xi_k x_{k,i}
$$
where $\xi_k$ is the intercept and regression coefficients for linear predictors. Meanwhile, the region-specific model assumes the effects of watershed area and branching probability differ between regions (i.e., interactions between region and watershed area/branching probability).

$$
\mu_i = \xi_0 + \xi_1~log_{10}~wa_i + \xi_2~log_{10}~p_{b,i} + \xi_3~region_i + \xi_4~log_{10}~WA_i \cdot region_i + \xi_5~log_{10}~P_{b,i} \cdot region_i + \sum_k \xi_k x_{k,i}
$$
We compared the performance of these competing models using the Bayes factor[@wagenmakersPracticalSolutionPervasive2007], which gives the relative strength of evidence in favor of the global model (M0) over the region-specific model (M1). We used Bayesian Information Criteria (BIC) of the linear models to approximate the Bayes factor[@wagenmakersPracticalSolutionPervasive2007]:

$$
Bayes~factor = exp(\Delta BIC_{10}/2)
$$
where $\Delta BIC_{10} = BIC(M1) - BIC(M0)$.
